Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-11-11 manuscript/supplemental_material.Rnw
fix placenta class analysis name
switch from barrier to intimacy

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=15pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=blue,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\usepackage{xspace}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<placental_classification,echo=FALSE,results="hide",cache.extra=file.info(dir("../data","placenta.*classification.*",full.names=TRUE))[,"mtime"]>>=
load("../data/placenta_classification")
load("../data/placenta_classification_p")
for (mf.bp in c("mf","bp")) {
    for (class in c("interdigitation","barrier","shape")) {
        load(paste0("../data/placenta_classification_go_",mf.bp,"_",class))
    }
}

@@

<<placenta_class_table,echo=FALSE,results="hide">>=
placenta_class_table <- function(class="shape") {
    subtable <- placenta.classification.p[factor==class & type == "aov",list(gene,p,fdr)]
    setnames(subtable,c("Gene","p","FDR"))
    subtable <- head(subtable[FDR <= 0.05,][order(p),],n=50)
    print(xtable(subtable,
                 align=c("c","c","r","r"),
                 display=c("s","s","g","g")
                 ),
          booktabs=TRUE,
          floating=FALSE,
          include.rownames=FALSE,
          math.style.exponents=TRUE,
          size="\\scriptsize")
}

@@

\begin{table}[h]
<<attachment_shape_significant,echo=FALSE,results="asis">>=
placenta_class_table("shape")
@ 
\caption{Top 50 genes which are associated with the shape of placenta attachment}
\label{tab:shape_genes}
\end{table}
\clearpage

@@

\begin{table}[h]
<<intimacy_significant,echo=FALSE,results="asis">>=
placenta_class_table("barrier")
@ 
\caption{Top 50 genes which are associated with the degree of placenta intimacy}
\label{tab:barrier_genes}
\end{table}

@@

@@

\begin{table}
<<pathway_analysis_class,echo=FALSE,results="asis",cache.extra=file.info(dir("../data","placenta_gsea.*",full.names=TRUE))[,"mtime"]>>=
pathway <- list()
for (class in c("shape","interdigitation","barrier")) {
    for (database in c("kegg","reactome","biocarta","netpath")) {
        load(paste0("../data/placenta_gsea_",class,"_",database))
        eval(parse(text=paste0("gsea.table <- placenta_gsea_",class,"_",database)))
        gsea.table[,Database:=gsub("^(.)","\\U\\1",database,perl=TRUE)]
        gsea.table[,Class:=gsub("^(.)","\\U\\1",class,perl=TRUE)]
        pathway[[paste0(class,"_",database)]] <-
            gsea.table
    }
}
placenta_class_gsea <-
    rbindlist(pathway)[fdr < 0.1,]
placenta_class_gsea[,Pathway:=gsub("^(KEGG|BIOCARTA|REACTOME) ","",
                                   gsub("_+"," ",
                                        Pathway))]
placenta_class_gsea[,Pathway:=gsub(".+\\%","",Pathway)]
setnames(placenta_class_gsea,"p_values","p")
setorder(placenta_class_gsea,"p")
print(xtable(placenta_class_gsea,
             display=c("s","s","g","g","s","s"),
             align=c("c","p{3in}","r","r","c","c")),
      math.style.exponents=TRUE,
      include.rownames=FALSE,
      booktabs=TRUE,
      floating=FALSE)      
@ 
\caption[Placenta Classification Pathway Enrichment]{Pathways with
  significant and suggestive enrichment ($\text{FDR}\le 0.1$) in genes
  associated with placenta morphology using the biocarta, reactome,
  and kegg pathway databases.}
\label{tab:core_placenta_pathway_analysis}
\end{table}

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=15pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=blue,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\usepackage{xspace}
% \usepackage[below]{placeins}
% \makeatletter
% \AtBeginDocument{%
%   \expandafter\renewcommand\expandafter\subsection\expandafter
%     {\expandafter\@fb@secFB\subsection}%
%   \newcommand\@fb@secFB{\FloatBarrier
%     \gdef\@fb@afterHHook{\@fb@topbarrier \gdef\@fb@afterHHook{}}}%
%   \g@addto@macro\@afterheading{\@fb@afterHHook}%
%   \gdef\@fb@afterHHook{}%
% }
% \makeatother
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<placental_classification,echo=FALSE,results="hide",cache.extra=file.info(dir("../data","placenta.*classification.*",full.names=TRUE))[,"mtime"]>>=
load("../data/placenta_classification")
load("../data/placenta_classification_p")
for (mf.bp in c("mf","bp")) {
    for (class in c("interdigitation","intimacy","shape")) {
        load(paste0("../data/placenta_classification_go_",mf.bp,"_",class))
    }
}

@@

<<placenta_class_table,echo=FALSE,results="hide">>=
placenta_class_table <- function(class="shape",analysis.type="aov") {
    subtable <- placenta.classification.p[factor==class & type == analysis.type,
                                          list(gene,p,fdr)]
    setnames(subtable,c("Gene","p","FDR"))
    subtable <- head(subtable[order(p),],n=50)
    top.type <- function(gene.name) {
        placenta.classification.p[gene==gene.name & factor==class
                                  & type =="glm",][order(-abs(coefficient))][1,level]
    }
    subtable[,`level_type`:=sapply(Gene,top.type)]
    setnames(subtable,"level_type",
             paste0("Most Significant ",
                    gsub("^(.)","\\U\\1",class,perl=TRUE)))
    print(xtable(subtable,
                 align=c("c","c","r","r","c"),
                 display=c("s","s","g","g","s")
                 ),
          booktabs=TRUE,
          floating=FALSE,
          include.rownames=FALSE,
          math.style.exponents=TRUE,
          size="\\scriptsize")
}

@@

\begin{table}
<<attachment_shape_significant,echo=FALSE,results="asis">>=
placenta_class_table("shape")
@ 
\caption{Top 50 genes which are associated with the shape of placenta attachment}
\label{tab:shape_genes}
\end{table}

@@

\begin{table}[h]
<<intimacy_significant,echo=FALSE,results="asis">>=
placenta_class_table("intimacy")
@ 
\caption{Top 50 genes which are associated with the degree of placenta intimacy}
\label{tab:intimacy_genes}
\end{table}

\begin{table}[h]
<<intimacy_significant_polr,echo=FALSE,results="asis">>=
placenta_class_table("intimacy",analysis.type="polr")
@ 
\caption{Top 50 genes which are associated with the degree of placenta intimacy; ordered analysis}
\label{tab:intimacy_genes_polr}
\end{table}

@@

\begin{table}[h]
<<interdigitation_significant_polr,echo=FALSE,results="asis">>=
placenta_class_table("interdigitation",analysis.type="polr")
@ 
\caption{Top 50 genes which are associated with the degree of maternal-fetal interdigitation; ordered analysis}
\label{tab:interdigitation_genes_polr}
\end{table}

@@

\begin{table}
  \small
<<pathway_analysis_class,echo=FALSE,results="asis",cache.extra=file.info(dir("../data","placenta_gsea.*",full.names=TRUE))[,"mtime"]>>=
pathway <- list()
for (class in c("shape","interdigitation","intimacy")) {
    for (database in c("kegg","reactome","biocarta","netpath")) {
        load(paste0("../data/placenta_gsea_",class,"_",database))
        eval(parse(text=paste0("gsea.table <- placenta_gsea_",class,"_",database)))
        gsea.table[,Database:=gsub("^(.)","\\U\\1",database,perl=TRUE)]
        gsea.table[,Class:=gsub("^(.)","\\U\\1",class,perl=TRUE)]
        pathway[[paste0(class,"_",database)]] <-
            gsea.table
    }
}
placenta_class_gsea <-
    rbindlist(pathway)[fdr < 0.2,]
placenta_class_gsea[,Pathway:=gsub("^(KEGG|BIOCARTA|REACTOME) ","",
                                   gsub("_+"," ",
                                        Pathway))]
placenta_class_gsea[,Pathway:=gsub(".+\\%","",Pathway)]
setnames(placenta_class_gsea,"p_values","p")
setorder(placenta_class_gsea,"p")
print(xtable(placenta_class_gsea,
             display=c("s","s","g","g","s","s"),
             align=c("c","p{3in}","r","r","c","c")),
      math.style.exponents=TRUE,
      include.rownames=FALSE,
      booktabs=TRUE,
      floating=FALSE)      
@ 
\caption[Placenta Classification Pathway Enrichment]{Pathways with
  significant and suggestive enrichment ($\text{FDR}\le 0.2$) in genes
  associated with placenta morphology using the biocarta, reactome,
  and kegg pathway databases.}
\label{tab:placenta_pathway_class_analysis}
\end{table}
